### 2019-12-15 proof-adaptive-optimality.tex

\section{Proof of Theorem~\ref{thm:adpative_lower_bound}}
\label{proof:thm:adpative_lower_bound}

We begin with two technical lemmas.
\begin{lem} \label{lem:bound_intervals}
  Let $f$ be a log-concave and symmetric density function
  for which Assumption~\ref{assump:failure_rate} holds. For any $x_1 \geq
  \ldots \geq x_n \in \R$,
  \begin{equation}
    \frac{ \left| \sum_{k=1}^n (-1)^{k+1} f(x_k) \right|^2 }{
      \left( \sum_{k=1}^n (-1)^{k+1} F(x_k) \right) \left(1- \sum_{k=1}^n (-1)^{k+1} F(x_k) \right) } 
    \leq  4f(0)^2.
    \label{eq:lem_bound_intervals}
  \end{equation}
\end{lem}
\begin{lem} \label{lem:fisher_bound}
  Let $X$ be a random variable with a symmetric, log-concave, and
  continuously differentiable density function $f(x)$ such that Assumption~1
  holds. For a Borel measurable set $A$, define
  \begin{equation*}
    B(x) \defeq \begin{cases} 1
      & \mbox{if} ~x \in A, \\
      -1 & \mbox{if}~ x \notin A.
    \end{cases}
  \end{equation*}
  Then the Fisher information of $B$ with respect to $\theta$ is bounded
  from above by $\eta(0)$.
\end{lem}

Lemma~\ref{lem:bound_intervals} is the special case $\delta = 0$ of
Lemma~\ref{lem:bound_intervals_delta} to come in
Section~\ref{sec:bound_intervals_delta}.  We now prove
Lemma~\ref{lem:fisher_bound}.

\begin{proof-of-lemma}[\ref{lem:fisher_bound}]  
  When $f$ is the normal density function, this lemma follows from
  \cite[Thm.~3]{Barnes2018}. The proof below is based on a different techique
  than in \cite{Barnes2018}, and is valid for any log-concave symmetric
  density satisfying Assumption~\ref{assump:failure_rate}. \\

  The Fisher information of $B$ with respect to $\theta$ is given by
  \begin{align}
    I_\theta & =  \ex{ \left( \frac{d}{d\theta} \log P\left( B | \theta \right) \right)^2 |\theta } \nonumber \\
    & = \frac{ \left(\frac{d}{d\theta} P(B=1|\theta) \right)^2}{P(B=1| \theta)} + \frac{ \left(\frac{d}{d\theta} P(B=-1|\theta) \right)^2} {P(B=-1| \theta)} \nonumber \\
    & =  \frac{ \left( \frac{d}{d\theta} \int_A f \left( x-\theta\right)dx \right)^2} { P(B=1| \theta) } + \frac{ \left( \frac{d}{d\theta}\int_A f \left( x-\theta \right)dx \right)^2} { P(B=-1| \theta) } \nonumber \\ 
    & \overset{(a)}{=} \frac{ \left( - \int_A f' \left( x-\theta \right)dx \right)^2} { P(B=1| \theta) } + \frac{ \left(- \int_A f' \left( x-\theta \right)dx \right)^2} { P(B=-1| \theta) } \nonumber \\ 
    & = \frac{\left( \int_A f'\left( x-\theta \right) dx \right)^2 }{  P(B=1 | \theta) \left(1-P(B=1|\theta) \right)  }, \nonumber \\
    & = \frac{\left( \int_A f'\left( x-\theta \right) dx \right) \left( \int_A f'\left( x-\theta \right) dx \right)}{ \left( \int_A f \left( x-\theta \right) dx \right)  \left(1- \int_A f \left( x-\theta \right) dx \right) }, \label{eq:lem_fisher_bound_proof1}
  \end{align}
  where differentiation under the integral sign in $(a)$ is possible since
  $f$ is log-concave and so a.e.\ differentiable
  (cf.~\cite{Bertsekas73}),
  with a.e.\ derivative $f'(x)$. Regularity of
  the Lebesgue measure implies that for any $\epsilon>0$, there exists a
  finite number $k$ of disjoint open intervals $I_1,\ldots I_k$ such that
  \begin{equation*}
    \int_{A\setminus \cup_{j=1}^k I_j }  dx < \epsilon,
  \end{equation*}
  which implies that for any $\epsilon' > 0$, the set $A$ in
  \eqref{eq:lem_fisher_bound_proof1} can be replaced by a finite union of
  disjoint intervals without increasing $I_\theta$ by more than
  $\epsilon'$. It is therefore enough to proceed in the proof assuming that
  $A$ is of the form
  \begin{equation*}
    A = \cup_{j=1}^k (a_j,b_j),
  \end{equation*}
  with $\infty \leq a_1 \leq \ldots a_k$, $b_1 \leq b_k \leq \infty$ and $a_j \leq b_j$ for $j=1,\ldots,k$. Under this assumption we have
  \begin{align*}
    \mathbb P(B_n=1| \theta) & = \sum_{j=1}^k \mathbb P\left(X_n \in (a_j,b_j) \right)  \\
    & = \sum_{j=1}^k \left( F \left(b_j-\theta\right) -  F \left(a_j-\theta\right)  \right),
  \end{align*}
  so we may rewrite Eq.~\eqref{eq:lem_fisher_bound_proof1} as
  \begin{align*}
    I_\theta & =   \frac { \left( \sum_{j=1}^{k} f \left(a_j-\theta \right) - f \left( b_j-\theta \right)  \right)^2 } 
    { \left( \sum_{j=1}^k F \left( b_j-\theta \right) - F \left( a_j-\theta \right)  \right) }  \nonumber \\
    & \times \frac {1} 
    {1- \left( \sum_{j=1}^k F \left(  b_j-\theta \right) - F \left( a_j-\theta \right)  \right) } 
  \end{align*}
  It follows from Lemma~\ref{lem:bound_intervals} that for any $\theta \in
  \R$ and any choice of the intervals' endpoints,
  \begin{equation*}
    I_\theta \le
    \max_{t \in \{a_j,b_j, j=1,\ldots,k\} } 4f(t)^2 \leq 4 f(0)^2, 
  \end{equation*}
  where the last transition is by Assumption~\ref{assump:failure_rate}. 
\end{proof-of-lemma}

We now finish the proof of Theorem~\ref{thm:adpative_lower_bound}.
To bound the Fisher information of any set of $n$ one-bit
messages with respect to $\theta$,
%% we first note that without loss of
%% each message $B_i$ can is of the form
%% \begin{equation}
%% B_i = \begin{cases}
%% X_i \in A_i & 1, \\
%% X_i \notin A_i & -1,
%% \end{cases} 
%% \end{equation}
%% where $A_i \subset \R$ is a Borel measurable set. 
%% %Indeed, any measurable function $(X_i) \in \{-1,1\}$ can be written in the form \eqref{eq:general_messages} with $A_i = B^{-1}(1)$.
we consider the conditional distribution $P({B_1,\ldots,B_n|\theta})$ of
$(B_1,\ldots,B_n)$ given $\theta$. We have
\begin{align*}
  P\left( B_1,\ldots,B_n | \theta \right) & =  \prod_{i=1}^n P\left(B_i | \theta, B_1,\ldots,B_{i-1} \right),
  %% \label{eq:adpt_lower_bound_proof:1}
\end{align*}
where $P\left(B_i =1 | \theta, B_1,\ldots,B_{i-1} \right) = \mathbb P\left(
X_i \in A_i\right)$, when $B_i = \indic{X_i \in A_i}$ by independence,
so that the Fisher information of $B_1,\ldots,B_n$ with
respect to $\theta$ is
\begin{align}
I_\theta(B_1,\ldots,B_N) = \sum_{i=1}^n I_\theta (B_i|B_1,\ldots,B_{i-1}),
\label{eq:fisher_information}
\end{align}
where $I_\theta (B_i|B_{i-1},\ldots,B_1)$ is the Fisher information of the
distribution of $B_i$ given $B_1,\ldots,B_{i-1}$.
Lemma~\ref{lem:fisher_bound} implies that $I_\theta
(B_i|B_{i-1},\ldots,B_1) \leq 4f(0)^2$. The Van Trees inequality
\cite{van2004detection, gill1995applications} now implies
\begin{align*}
\ex{ \left( \theta_n - \theta \right)^2} &  \geq \frac{1}{ \ex{ I_\theta(B_1,\ldots,B_n)} + I_0} \\
& = \frac{1}{ \sum_{i=1}^n I_\theta (B_i | B^{i-1} ) + I_0} \\
& \geq \frac{1}{ 4f^2(0) n + I_0}
\end{align*}
as desired.

\subsection{An Isoperimetric Lemma}
\label{sec:bound_intervals_delta}

The following lemma is essential to the the proofs
of Theorems~\ref{thm:adpative_lower_bound}, \ref{thm:LAN1}, and
\ref{thm:non_existence}.

\begin{lem}
  \label{lem:bound_intervals_delta}
  Let $f$ be a log-concave and symmetric density
  function. Let $\delta\geq 0$. Assume that the function
  \begin{equation*}
    \eta_\delta(x) \triangleq  \eta^{1+\delta}(x)/f^\delta(x)
    = \frac{  \left( f(x) \right)^{2+\delta}}{\left(F(x)(1-F(x))\right)^{1+\delta}}
  \end{equation*}
  is non-increasing in $|x|$. Then for any $x_1 \ge \ldots \ge x_n \in \R^n$,
  \begin{equation}
    \frac{ \left| \sum_{i=1}^n (-1)^{i+1} f(x_i) \right|^{2+\delta} } {\left|
      \sum_{i=1}^n (-1)^{i+1} F(x_i) \right|^{1+\delta} \left|1- \sum_{k=1}^n
      (-1)^{i+1} F(x_i) \right|^{1+\delta} } \leq \max_{i} \eta_\delta(x_i).
    \label{eq:lem_bound_intervals_delta}
  \end{equation}
\end{lem}
In particular, 
\begin{equation*}
\frac{ \left| \sum_{i=1}^n (-1)^{i+1} f(x_i) \right|^{2+\delta} }
{\left| \sum_{i=1}^n (-1)^{i+1} F(x_i) \right|^{1+\delta} \left|1- \sum_{i=1}^n (-1)^{i+1} F(x_i) \right|^{1+\delta} } 
\leq \eta_\delta(0) = 4^{1+\delta} f^{2+\delta}(0).
\end{equation*}

\begin{proof-of-lemma}[\ref{lem:bound_intervals_delta}]
  Denote 
  \begin{equation*}
    \delta_n(x_1,\ldots,x_n) \triangleq \sum_{i=1}^{n} s_i f(x_i),
  \end{equation*}
  \begin{equation*}
    \Delta_n(x_1,\ldots,x_n) \triangleq  \sum_{i=1}^n s_i F(x_i), 
  \end{equation*}
  where $s_i \triangleq (-1)^{i+1}$. We use induction on $n \in \N$ to show that 
  \begin{align}
    \label{eq:lemm:interval_bounds:to_show}
    \frac{ \left| \delta_n(x_1,\ldots,x_n) \right|^{2+\delta}} 
         {\left|\Delta_n(x_1,\ldots,x_n)\left(1- \Delta_n(x_1,\ldots,x_n) \right) \right|^{1+\delta} } \leq \max_{i}\eta_{\delta}(x_i).
  \end{align}
  Since 
  \begin{equation*}
    \eta_\delta(x) =  \frac{  \left|\delta_1(x) \right|^{2+\delta}}{\left|\Delta_1(x)
      (1-\Delta_1(x)) \right|^{1+\delta}}, 
  \end{equation*}
  The case $n=1$ is trivial.  
  %follows from the assumption that $\eta_\delta(x)$ is non-increasing in $|x|$. 
  Assume that \eqref{eq:lemm:interval_bounds:to_show} holds for all integers up to $n = N$ and for any $x_1 \ge \ldots \ge x_N$. Consider the case $n = N+1$. Let $i^*$ be the index such that $x_{i^*}$ has minimal absolute value among $x_1,\ldots,x_N$. The assumption on $\eta_\delta(x)$ implies that
  \begin{equation*}
    \eta_\delta(x_{i^*}) = \max_i \eta_\delta(x_i).
  \end{equation*}
  Since the LHS of \eqref{eq:lem_bound_intervals_delta} is invariant to a sign flip of all $x_1,\ldots,x_{N+1}$, we may assume that $x_{i^*}$ is positive without loss of generality. 
  %For simplicity, we also assume that $s_{i^*} = 1$ and note that the case $s_{i^*} = -1$ is obtained using identical arguments. 
  Set $x^* = x_{i^*}$ and let $k=i^*-1$. In what follows, variables with subscript of non-positive index are ignored in summations and in lists of arguments to functions. Consider the function
  \begin{align}
    & g(y_1,\ldots,y_N) \triangleq g(y_1,\ldots,y_N|x^*,k) \\
    &  \triangleq  \frac{\left| \delta_{N+1}(y_1,\ldots,y_k,x^*,y_{k+1}\ldots,y_N) )\right|^{2+\delta}}{
      \left|\Delta_{N+1}(y_1,\ldots,y_k,x_{i^*},y_{k+1}\ldots,y_N)  (1 -\Delta_{N+1}(y_1,\ldots,y_k,x^*,y_{k+1}\ldots,y_N)  \right|^{1+\delta}
    } \label{eq:g_def}.
  \end{align}
  The LHS of \eqref{eq:lemm:interval_bounds:to_show} is obtained by taking $y_i=x_{k_i}$ where $k_i$ is the $i$th element in $\{1,\ldots,N+1\}\setminus \{i^*\}$. It is therefore enough to prove that 
  \begin{equation*}
    \max_{(y_1,\ldots,y_N) \in A_N} g(y_1,\ldots,y_N) \leq \eta_{\delta}(x^*),
  \end{equation*}
  where 
  \begin{equation*}
    A_N(x^*,k) \triangleq \left\{ (y_1,\ldots,y_N) \in \R^N\, : \, y_1 \ge y_k \ge x^* \ge -x^* \ge y_{k+1} \ldots \ge y_N
    \right\}.
  \end{equation*}
  %
  Since $f(x)$ is log-concave and symmetric, we may write $f(x) = e^{c(x)}$
  where $c(x)$ is concave, symmetric, and superdifferentiable on the interior
  of its domain with
  supergradient set $\partial c(x) = \{v \in \R \mid c(y) \le c(x) + v (y - x) ~
  \mbox{for~all}~ y\}$; $c$ is also
  differentiable a.e.\ with derivative
  \begin{equation*}
    c'(x) \triangleq \frac{f'(x)}{f(x)}
  \end{equation*}
  (when it exists), and we otherwise simply treat $f'(x) / f(x) = c'(x) \in
  \partial c(x)$ as an arbitrary element of the superdifferential.  The
  supergradient sets $\partial c(x)$ are increasing, in that $v_0 \in
  \partial c(x_0)$ and $v_1 \in \partial c(x_1)$ implies that $(v_1 - v_0)
  (x_1 - x_0) \le 0$.  We first prove the lemma under the assumption that
  $c$ is strictly concave, or, equivalently, that $v_i \in \partial c(x_i)$
  implies that $(v_1 - v_0)(x_1 - x_0) < 0$ whenever $x_1 \neq x_0$; that
  is, $c'$ is strictly decreasing.

  The maximal value of $g(y_1,\ldots,y_N)$ is attained for the same
  $(y_1,\ldots,y_N) \in A_N(x^*,k)$ that maximizes
  \begin{align*}
    \log(g)(y_1,\ldots, y_N) =  (2+\delta) \log \left( \delta_N  \right)  - (1+\delta) \log \left( \Delta_N  \right) - (1+\delta) \log \left(1 - \Delta_N \right),
  \end{align*}
  where in the last display and henceforth we suppress the arguments
  $y_1,\ldots,y_k,x^*,y_{k+1},\ldots, y_N$ of the functions $\delta_N$ and
  $\Delta_N$. (We may take the compactification of $\R$ as $[-\infty, \infty]$.)
  %
  Within the interior of $A_N(x^*,k)$, all three expressions in
  \eqref{eq:g_def} within an absolute value are positive. It follows that
  partial derivative of $\log(g)(y_1,\ldots,y_N)$ with respect to $y_i$ within
  the interior of $A_N(x^*,k)$ is
  \begin{equation*}
    \frac{\partial \log(g)}{\partial y_i} = \frac{(2+\delta) s_i
      f'(x_i)}{\delta_N } -\frac{(1+\delta) s_i f(x_i)}{\Delta_N } +
    \frac{(1+\delta)s_i f(y_i)}{1-\Delta_N }.
  \end{equation*}
  We conclude that the gradient of $\log(g)$ vanishes if and only if
  \begin{equation}
    \label{eq:gradient_zero}
    c'(y_i) = \frac{f'(y_i)}{f(y_i)} = \frac{1+\delta}{2+\delta} \frac{\delta_N}{2} \left( \frac{1}{\Delta_N} - \frac{1}{1-\Delta_N } \right),\quad i=1,\ldots,N.
  \end{equation}
  %
  Since we assumed that $\partial c(x)$ is injective,
  equality~\eqref{eq:gradient_zero} holds if and only if $y_1 = \ldots
  = y_N$. In this case, $g(y_1,\ldots,y_N) = \eta_\delta(x_{i^*})$ if $N$ is
  even. If $N$ is odd and $y_1 = \ldots = y_N > x^*$, then
  \begin{align*}
    & g(y_1,\ldots,y_N) = \frac{\left| f(y_1)-f(x_{i^*})  \right|^{2+\delta}} { 
      \left|  F(y_1) - F(x_{i^*}) \right|^{1+ \delta} 
      \left| 1 - (F(y_1) -F(x_{i^*})) \right|^{1+ \delta} } 
  \end{align*} 
  which is bounded from from above by $\eta_\delta(x_{i^*})$ by the induction hypothesis. The case where $N$ is odd and $-x^* \leq y_1 = \ldots = y_N$ is similar. 
  %
  We now consider the possibility that the maximum of $g(y_1,\ldots,y_N)$ is attained at the boundaries of $A_N(x^*,k)$. At boundary points for which $y_i = y_{i+1}$ for some $i$, the contribution of $y_i$ and $y_{i+1}$ to $g(y_1,\ldots,y_N)$ is zero and the induction assumption for $n=N-1$ implies that 
  \begin{equation*}
    g(y_1,\ldots,y_N) \leq \eta_{\delta}(x^*)
  \end{equation*}
  The remaining boundary points of $A_N(x^*,k)$ are covered by the following cases:
  \begin{itemize}
  \item[(1)]  $y_N \to -\infty$. 
  \item[(2)] $y_1 \to \infty$.
  \item[(3)] $y_k = x_{i^*}$.
  \item[(4)] $y_{k+1} = -x_{i^*}$. 
  \end{itemize}
  For case (1), 
  \begin{align*}
    g(y_1,\ldots,y_N) \to \frac{ \left| \sum_{i=1}^{k} s_i f(y_i) + s_{i^*} f(x_{i^*}) - \sum_{i=k+1}^{N-1} s_i f(y_i) \right|^{2+\delta}} 
    {\left| \sum_{i=1}^{k} s_i F(y_i) + s_{i^*} F(x_{i^*}) - \sum_{i=k+1}^{N-1} s_i F(y_i) \right|^{1+\delta}\left|1- \left( \sum_{i=1}^{k} s_i F(y_i) + s_{i^*} F(x_{i^*}) - \sum_{i=k+1}^{N-1} s_i F(y_i) \right)  \right|^{1+\delta} },
  \end{align*}
  which is smaller than $\eta_\delta(x_{i^*})$ by the induction hypothesis. Similarly, under case (2),
  \begin{align*}
    & g(y_1,\ldots,y_N) \to 
    \frac{ \left| \sum_{i=2}^{k} s_i f(y_i) + s_{i^*} f(x_{i^*}) - \sum_{i=k+1}^{N} s_i f(y_i) \right|^{2+\delta}} 
         {\left| 1+ \sum_{i=2}^{k} s_i F(y_i) + s_{i^*} F(x_{i^*}) - \sum_{i=k+1}^{N} s_i F(y_i) \right|^{1+\delta}\left|-\left( \sum_{i=2}^{k} s_i F(y_i) + s_{i^*} F(x_{i^*}) - \sum_{i=k+1}^{N} s_i F(y_i) \right)  \right|^{1+\delta} },
         \\
         & = \frac{ \left| -\sum_{i=2}^{k} s_i f(y_i) - s_{i^*} f(x_{i^*}) + \sum_{i=k+1}^{N} s_i f(y_i) \right|^{2+\delta}} 
           { \left|1 - \left(-\sum_{i=2}^{k} s_i F(y_i) - s_{i^*} F(x_{i^*}) + \sum_{i=k+1}^{N} s_i F(y_i) \right) \right|^{1+\delta} 
             \left|-\sum_{i=2}^{k} s_i F(y_i) - s_{i^*} F(x_{i^*}) + \sum_{i=k+1}^{N} s_i F(y_i) \right|^{1+\delta} }
  \end{align*}
  which is smaller than $\eta_{\delta}(x_{i^*})$ by the induction hypothesis. Under case (3), the terms in $\delta_N$ and $\Delta_N$ corresponding to $y_k$ and $x_{i^*}$ cancel each other. As a result,  $g(y_1,\ldots,y_N)$ reduces to an expression with $n=N-1$ variables hence this case is handled by the induction hypothesis. 
  %
  Finally, under case (4), set 
  \begin{equation*}
    d \triangleq s_k F(-x^*) + s_{i^*} F(x^*) = s_{i^*}\left(1-2F(-x^*) \right), 
  \end{equation*}
  \begin{equation*}
    \sigma \triangleq \sum_{i=1}^{k-1} s_i f(y_i) - \sum_{i=k+1}^{N} s_i f(y_i). 
  \end{equation*}
  and 
  \begin{equation*}
    \Sigma \triangleq \sum_{i=1}^{k-1} s_i F(y_i) - \sum_{i=k+1}^{N} s_i F(y_i). 
  \end{equation*}
  We have
  \begin{align*}
    & g(y_1,\ldots,y_N) =\nonumber  \\ 
    & =
    \frac{ \left| \sum_{i=1}^{k-1} s_i f(y_i) - \sum_{i=k+1}^{N} s_i f(y_i) \right|^{2+\delta}} 
         {\left| \sum_{i=1}^{k-1} s_i F(y_i) + d(x^*) - \sum_{i=k+1}^{N} s_i F(y_i) \right|^{1+\delta} \left|1- \sum_{i=1}^{k-1} s_i F(y_i) - d(x^*) + \sum_{i=k+1}^{N} s_i F(y_i)   \right|^{1+\delta} }, \\
         & = \frac{ \left| \sigma \right|^{2+\delta}} 
         {\left| \Sigma+d \right|^{1+\delta} \left|1- \Sigma - d \right|^{1+\delta} }  =  
         \frac{ \left| \sigma \right|^{2+\delta}} 
              {\left| \Sigma \right|^{1+\delta} \left|1- \Sigma \right|^{1+\delta} }  \left| \frac{\Sigma(1-\Sigma) } { \Sigma(1-\Sigma) + d(1-2\Sigma)-d^2} \right|^{1+\delta}. 
  \end{align*}
  By the induction hypothesis,
  \begin{equation*}
    \frac{ \left| \sigma \right|^{2+\delta}} 
         {\left| \Sigma \right|^{1+\delta} \left|1- \Sigma \right|^{1+\delta} } \leq \eta_\delta(x^*), 
  \end{equation*}
  hence it is left to show that 
  \begin{equation*}
    \frac{\Sigma(1-\Sigma) } { \Sigma(1-\Sigma) + d(1-2\Sigma)-d^2} \leq 1.
  \end{equation*}
  Whenever $d>0$, 
  \begin{align*}
    & \frac{ \Sigma(1-\Sigma) + d(1-2\Sigma)-d^2} {\Sigma(1-\Sigma)} \geq 1 \Leftrightarrow  1-2\Sigma \geq d, 
  \end{align*}
  while for $d<0$,
  \begin{align*}
    & \frac{ \Sigma(1-\Sigma) + d(1-2\Sigma)-d^2} {\Sigma(1-\Sigma)} \geq 1 \Leftrightarrow  1-2\Sigma \leq d. 
  \end{align*}
  Therefore, it is enough to show that $\Sigma \leq F(-x^*)$ if $s_{i^*}=1$ and 
  $\Sigma \geq F(-x^*)$ if $s_{i^*}=-1$. 
  %
  Indeed, if $s_{i^*}=1$, then $s_{k+1}=-1$ and monotonicity of $F(x)$ implies that 
  \begin{equation*}
    \Sigma + d \leq F(y_1) - F(y_k) + F(x^*) - F(-x^*) + F(y_{k+2}) - F(y_N), 
  \end{equation*}
  and hence
  \begin{equation*}
    \Sigma \leq 1-F(x^*) = F(-x^*). 
  \end{equation*}
  Similarly, if $s_{i^*}=-1$ then 
  \begin{equation*}
    1 - \Sigma \leq 1 -  F(-x^*).
  \end{equation*}
  This conclude the proof in the case where $c'(x)$ is an injection.

  In the case where $c(x)$ is not strictly concave, so that $c'$ does not
  strictly decrease, we approximate $c$ using another concave symmetric
  function with decreasing derivative. We assume w.l.o.g.\ that $c(0) = 0$
  maximizes $c$.  For $\alpha>0$ consider the function $f_\alpha(x) =
  \kappa(\alpha) e^{-|c(x)|^{1+\alpha}}$, where $\kappa(\alpha)$ normalizes
  $f_\alpha$.  Then $c_\alpha(x)$ is concave, symmetric, and
  a.e.\ differentiable with
  \begin{equation*}
    c_\alpha'(x) \triangleq \frac{f'_\alpha(x)}{f_\alpha(x)} = (1+\alpha)|c(x)|^{\alpha} c'(x). 
  \end{equation*}
  Now $c_\alpha'(x)$ is non-increasing since it is the derivative of a concave
  function. Furthermore, since $c(x)$ is non-constant on any interval and
  $c'(x)$ is non-increasing, $c_\alpha'(x)$ is non-constant on any interval
  hence an injection. It follows from the first part of the proof that, for
  any $\alpha>0$,
  \begin{align}
    \label{eq:proof:lem:bound_intervals}
    \frac{(\delta_{n,\alpha})^2}{\Delta_{n,\alpha}(1-\Delta_{n,\alpha})} \leq \max_i \eta_{\alpha}(x_i),
  \end{align}
  where 
  \begin{equation*}
    \delta_{n,\alpha} \triangleq  \sum_{k=1}^{n} (-1)^{k+1} f_{\alpha}(x_k),
  \end{equation*}
  \begin{equation*}
    \Delta_{n,\alpha} \triangleq \sum_{k=1}^n (-1)^{k+1} F_{\alpha}(x_k), 
  \end{equation*}
  and 
  \begin{equation*}
    \eta_{\delta,\alpha}(x) \triangleq \frac{(f_\alpha(x))^{2+\delta}}{\left(F_{\alpha}(x)(1-F(x)) \right)^{1+\delta}}. 
  \end{equation*}
  The proof is completed by noting that 
  \begin{align*}
    \lim_{\alpha \to 0} \frac{(\delta_{n,\alpha})^{2+\delta} }{ \left(\Delta_{n,\alpha}(1-\Delta_{n,\alpha})\right)^{1+\delta}}  = \frac{(\delta_{n})^{2+\delta }}{\left(\Delta_{n}(1-\Delta_{n}) \right)^{1+\delta}},  
  \end{align*}
  and, since the maximum is over a finite set,
  \begin{align*}
    \lim_{\alpha \to 0}  \max_i \eta_{\delta,\alpha}(x_i)  = \max_i\eta_\delta(x_i).
  \end{align*}

  %Our goal is to construct an approximation to $f(x)$ using a log-concave, symmetric, and differentiable function $\hat{f}(x)$ such that $\hat{f}'(x)/\hat{f}(x)$ is strictly decreasing. We achieve such an approximation by modifying $f(x)$ over intervals over which $c'(x)$ is a constant. Assume first $c'(x)$ is only a constant over an interval $(a_1,b_1)\subset \R$ containing $x_1>0$ and on the mirror image of this interval around $x=0$. Over this interval, $f(x)$ is necessarily of the form $f(x) = e^{-c_1|x| + d_1}$ for some $c_1\geq 0$ and $d_1$. % and For $\alpha>0$, consider
  %\begin{equation*}
  %F_1(x,\alpha) \triangleq  F(a_1) + \frac{1}{c_1} e^{-c_1 (|x|^{1+\alpha} +d_1}, \quad x \in (a_1,b_1),
  %\end{equation*}
  %and 
  %\begin{equation*}
  %f_1(x,\alpha) \triangleq  F_1'(x,\alpha) =  -\sgn(x)(1+\alpha)|x|^\alpha e^{-c_1 |x|^{1+\alpha} + d_1}, \quad x \in (a_1,b_1).
  %\end{equation*}

\end{proof-of-lemma}