### 2019-12-15 proof-lan-distributed.tex

% -*- Mode: latex -*- %

\section{Proof of Theorem~\ref{theorem:non-adaptive-minimax}}

We follow a similar outline to the optimality results we establish
in the proof of Theorem~\ref{thm:sgd}\eqref{item:sgd-regular} in
Sec.~\ref{sec:proof-sgd-regular}.
Roughly, we establish that the family $P_\theta$ of distributions
on the bits $B_i$ is locally asymptotically normal
(Definition~\ref{definition:lan}) via a quadratic
mean differentiability argument. After this, the result
follows by standard local asymptotic minimax theory.

\begin{lemma}
  \label{lemma:lan-bits}
  Let Assumption~\ref{assumption:detection-regions}\eqref{item:density-nice}
  and~\eqref{item:finite-intervals} hold, and let
  $B_i = \indic{X_i \in A_i}$.
  Let $h_n \to h \in \R$. Then for any $\theta \in \mbox{int}\Theta$,
  \begin{equation*}
    \sum_{i = 1}^n \log \frac{P_{\theta + h_n/\sqrt{n}}(B_i)}{
      P_\theta(B_i)}
    = \frac{h}{\sqrt{n}}
    \sum_{i = 1}^n \score_\theta(B_i)
    - \frac{h}{4n} \sum_{i = 1}^n \var(\score_\theta(B_i))
    - \frac{h}{4n} \sum_{i = 1}^n \score_\theta(B_i)^2
    + o_P(1).
  \end{equation*}
  If additionally
  Assumption~\ref{assumption:detection-regions}\eqref{item:limit-variance}
  holds, then
  \begin{equation*}
    \sum_{i = 1}^n \log \frac{P_{\theta + h_n/\sqrt{n}}(B_i)}{
      P_\theta(B_i)}
    = \frac{h}{\sqrt{n}}
    \sum_{i = 1}^n \score_\theta(B_i)
    - \frac{h^2}{2} \kappa(\theta) + o_P(1).
  \end{equation*}
\end{lemma}
\noindent
The proof of Lemma~\ref{lemma:lan-bits} is quite technical,
so we defer it to Section~\ref{sec:proof-lan-bits}.

With this lemma, it is not too challenging to demonstrate the local
asymptotic normality (Definition~\ref{definition:lan}) of the family
$\{P_\theta\}$. Indeed, the assumption that $f$ is Lipschitz guarantees that
$f(z) \le \sqrt{\lip(f)}$ for all $z$, and thus we obtain that
$|\dPtheta(A_n)| \le K \sqrt{\lip(f)}$ for all $n$, whence $\sup_n
\max\{|\score_\theta(A_n)|, |\score_\theta(A_n^c)|\} \le C < \infty$ for
some $C$. Thus, the strong law of large numbers
gives that
$\frac{1}{n} \sum_{i = 1}^n \score_\theta(B_i)
\cas 0$ as $\E_\theta[\score_\theta(B_i)] = 0$, and Lyapunov's central
limit theorem applies to give
\begin{equation*}
  \frac{1}{\sqrt{n}} \sum_{i = 1}^n \score_\theta(B_i)
  \cd \normal\left(0, \kappa(\theta)\right)
\end{equation*}
under
Assumption~\ref{assumption:detection-regions}\eqref{item:limit-variance},
so that the family $\{P_\theta\}$ is locally asymptotically normal
(Def.~\ref{definition:lan}).

We now recall the familiar H\'{a}jek-Le-Cam local asymptotic minimax
result~\cite[Thm.~8.11]{VanDerVaart98}: if the family
$\{P_\theta\}$ is LAN with precision $\kappa(\theta)$, then
\begin{equation*}
  \liminf_{c \to \infty} \liminf_n \sup_{\norm{\tau - \theta} \le
    c / \sqrt{n}} \E_\tau\left[L(\sqrt{n}(\theta_n - \tau))\right]
  \ge \E[L(Z / \sqrt{\kappa(\theta)})]
\end{equation*}
for any symmetric quasi-convex loss $L$, where $Z \sim normal(0, 1)$.
This immediately gives Theorem~\ref{theorem:non-adaptive-minimax}.

\subsection{Proof of Lemma~\ref{lemma:lan-bits}}
\label{sec:proof-lan-bits}

Our proof follows that of \cite[Thm.~7.2]{VanDerVaart98} closely.
The first step is to note that
if $\lip(f)$ denotes the Lipschitz constant of $f$, then for any
$h \in \R$ and $A = \cup_{i = 1}^K \{[t_i^-, t_i^+]\}$, we have
\begin{equation}
  \label{eqn:expansion-dPtheta}
  P_{\theta + h}(A) = P_\theta(A) + \dPtheta(A)
  \pm \frac{K \cdot \lip(f)}{2}  h^2.
\end{equation}
Indeed, for any interval $[t_0, t_1]$ we immediately have
\begin{align*}
  P_{\theta + h}([t_0, t_1])
  & = \int_{t_0}^{t_1} f(z - \theta - h) dz
  = \int_{t_0}^{t_1} \left(f(z - \theta)
  + \int_0^h f'(z - \theta - v) dv\right) dz \\
  & = P_\theta([t_0, t_1])
  + \int_0^h \int_{t_0}^{t_1} f'(z - \theta - v) dz dv \\
  & = P_\theta([t_0, t_1])
  + \int_0^h \left[f(t_1 - \theta - v) - f(t_0 - \theta - v)\right] dv \\
  & = P_\theta([t_0, t_1])
  + \left(f(t_1 - \theta) - f(t_0 - \theta)\right) h
  \pm \int_0^h \lip(f) |v| dv \\
  & = P_\theta([t_0, t_1]) + \dPtheta([t_0, t_1])
  \pm \frac{\lip(f)}{2} h^2,
\end{align*}
giving the result~\eqref{eqn:expansion-dPtheta}.

Now we demonstrate a type of uniform quadratic mean differentiability
(Definition~\ref{definition:qmd}) for such sets
$A$. First, note that by a Taylor approximation and concavity
of $\sqrt{\cdot}$, we have
\begin{equation*}
  \sqrt{a} + \frac{b}{2 \sqrt{a}} -
  \frac{b^2}{4\sqrt{a}}
  \le \sqrt{a + b} \le \sqrt{a}
  + \frac{b}{2 \sqrt{a}}
\end{equation*}
for any $a > 0$ and $|b| \le 3a/4$. Consequently,
recalling that $\score_\theta(A) = \dPtheta(A) / P_\theta(A)$,
for any $h \in \R$ the expansion~\eqref{eqn:expansion-dPtheta} yields
\begin{equation*}
  \left(\sqrt{P_{\theta + h}(A)} -
  \sqrt{P_\theta(A)} - \half h \score_\theta(A) \sqrt{P_\theta(A)}
  \right)^2
  \le
  \left(
  \frac{K \lip(f)}{2 \sqrt{P_\theta(A)}} h^2
  + \frac{(|\dPtheta(A) h| + h^2 \lip(f))^2}{P_\theta(A)^{3/2}}
  \right)^2.
\end{equation*}
Now, note that for any Lipschitz-continuous $f$, that
$1 = \int f(z) dz$ and $f \ge 0$ yield that
$f(z) \le \sqrt{\lip(f)}$ for all $z \in \R$. Thus, under
Assumption~\ref{assumption:detection-regions}\eqref{item:finite-intervals},
there exists a finite constant $C < \infty$ such that
\begin{subequations}
  \label{eqn:h-fourth}
  \begin{align}
    \nonumber \left(\sqrt{P_{\theta + h}(A)} -
    \sqrt{P_\theta(A)} - \half
    h \score_\theta(A) \sqrt{P_\theta(A)}\right)^2
    & \le
    \left(\frac{h^2 k \cdot \lip(f)}{2 \sqrt{P_\theta(A)}}
    + \frac{(|\dPtheta(A) h| + k h^2 \lip(f))^2}{
      P_\theta(A)^{3/2}}\right)^2 \\
    & \le \frac{C}{P_\theta(A)} \left[k^2 \lip^2(f)
      + \score_\theta(A)^2
      + \frac{k^4 h^4 \lip^4(f)}{P_\theta(A)^2}
      \right] \cdot h^4
  \end{align}
  where $C < \infty$ is a numerical constant, valid whenever $|\dPtheta(A) h|
  \le P_\theta(A) / 4$ and $k h^2 \lip^4(f) \le P_\theta(A) / 4$,
  and similarly, we have
  \begin{equation}
    \left(\sqrt{P_{\theta + h}(A^c)} -
    \sqrt{P_\theta(A^c)} - \half
    h \score_\theta(A^c) \sqrt{P_\theta(A^c)}\right)^2
    \le \frac{C}{P_\theta(A^c)}
    \left[k^2 \lip^2(f)
      + \score_\theta(A^c)^2
      + \frac{k^4 h^4 \lip^4(f)}{P_\theta(A^c)^2}
      \right] \cdot h^4.
  \end{equation}
\end{subequations}
That is, the family $\{P_\theta\}$ with bit observations $B_n$ satisfies a
uniform type of quadratic-mean differentiability
(Def.~\ref{definition:qmd}).

Now, for shorthand, define $P_n = P_{\theta + h_n / \sqrt{n}}$ and $P =
P_\theta$, and let $p_n, p$ be shorthand for the p.m.f.s of the two
distributions.  For the sets $A_i$ we recall that $B_i = \indic{X_i \in
  A_i}$.  The random variables
\begin{equation*}
  W_{n,i} \defeq 2 \left(\sqrt{\frac{p_n}{p}}(B_i) - 1\right)
\end{equation*}
are with $P$-probability 1 well-defined, and by the
inequalities~\eqref{eqn:h-fourth}, we have
that
\begin{align*}
  \var\left(W_{n,i} - \frac{h_n}{\sqrt{n}} \score_\theta(B_i)\right)
  & \le 
  C \frac{k^2 \lip^2(f) + \score_\theta(A_i)^2
  + \score_\theta(A_i^c)^2}{P_\theta(A_i) P_\theta(A_i^c)}
  \cdot \frac{h_n^4}{n^2}
  + C \frac{k^4 \lip^4(f)}{
    P_\theta(A_i)^3 P_\theta(A_i^c)^3}
  \frac{h_n^8}{n^4} \\
  & \le 
  C \frac{k^2 \lip^2(f) + \score_\theta(A_i)^2
  + \score_\theta(A_i^c)^2}{P_\theta(A_i) P_\theta(A_i^c)}
  \cdot \frac{h^4}{n^2}
  + C \frac{k^4 \lip^4(f)}{
    P_\theta(A_i)^3 P_\theta(A_i^c)^3}
  \frac{h^8}{n^4}
\end{align*}
whenever
\begin{equation*}
  \frac{h}{\sqrt{n}} \max\{\score_\theta(A_i),
  \score_\theta(A_i^c)\}
  \le \frac{1}{4}
  ~~ \mbox{and} ~~
  \frac{h^2}{n} \lip^2(f)
  \le \frac{\min\{P_\theta(A_i), P_\theta(A_i^c)\}}{4}
\end{equation*}
In particular, using the uniform boundedness of $0 < P_\theta(A_i) < 1$,
we see that unifomly in $i, n$ we have
\begin{equation*}
  \var\left(W_{n,i} - \frac{h}{\sqrt{n}} \score_\theta(B_i)\right)
  = O\left(\frac{h^4}{n^2}\right)
  + O(1) \E[\score_\theta(B_i)^2]
  \frac{(h_n - h)^2}{n},
\end{equation*}
while $\E[\score_\theta(B_i)] = 0$. Additionally, defining $\mu_i$ to be the
induced counting measure on $B_i = \indic{X_i \in A_i}$,
\begin{align*}
  \sum_{i = 1}^n \E[W_{n,i}]
  & = 2\sum_{i = 1}^n
  \left(\int \sqrt{p_n(b)} \sqrt{p(b)} d\mu_i(b) - 1 \right)
  = -\sum_{i = 1}^n \int \left(\sqrt{p_n(b)} - \sqrt{p(b)}\right)^2
  d\mu_i(b) \\
  & = -\frac{h_n^2}{4 n} \sum_{i = 1}^n \E[\score_\theta(B_i)^2]
  - \sum_{i = 1}^n \int \left(\sqrt{p_n(b)} - \sqrt{p(b)}
  - \frac{h_n}{\sqrt{n}} \score_\theta(b) \sqrt{p(b)}\right)^2 d\mu_i(b) \\
  & \qquad \qquad ~
  - \sum_{i = 1}^n \int \left(\sqrt{p_n(b)} - \sqrt{p(b)}
  - \frac{h_n}{\sqrt{n}} \score_\theta(b) \sqrt{p(b)}\right)
  \frac{h_n}{\sqrt{n}} \score_\theta(b) \sqrt{p(b)}d\mu_i(b) \\
  & = -\frac{h^2}{4n} \sum_{i = 1}^n \E[\score_\theta(B_i)^2]
  - O(h^4 / n) - O(h^2 / n^{1/2})
\end{align*}
uniformly in $h$.
Therefore, we obtain
\begin{equation*}
  \sum_{i = 1}^n W_{n,i}
  = \sum_{i = 1}^n \left(W_{n,i} - \frac{h_n}{\sqrt{n}} \score_\theta(B_i)\right)
  + \frac{h_n}{\sqrt{n}} \sum_{i = 1}^n \score_\theta(B_i)
  = -\frac{h^2}{4n} \sum_{i = 1}^n \E[\score_\theta(B_i)^2]
  + \frac{h}{\sqrt{n}} \sum_{i = 1}^n \score_\theta(B_i)
  + o_P(1),
\end{equation*}
where we have used that $h_n \to h$.

Now, we write the log-likelihood ratio. We have
\begin{align*}
  \sum_{i = 1}^n \log \frac{p_n(B_i)}{p(B_i)}
  & = 2 \sum_{i = 1}^n \log\left(1 + \half W_{n,i}\right) \\
  & = \sum_{i = 1}^n W_{n,i}
  - \frac{1}{4} \sum_{i = 1}^n W_{n,i}^2
  + \half \sum_{i = 1}^n W_{n,i}^2 R(W_{n,i})
\end{align*}
where the remainder $|R(W_{n,i})| \le |W_{n,i}|$ for $|W_{n,i}| \le 1$.
In particular, using the Taylor approximation to $\sqrt{\cdot}$ and
that $\inf_i P_\theta(A_i) > 0$, we have
$\max_i W_{n,i} \cas 0$, so that
\begin{equation}
  \label{eqn:almost-at-the-end}
  \sum_{i = 1}^n \log \frac{p_n(B_i)}{p(B_i)}
  = \frac{h}{\sqrt{n}} \sum_{i = 1}^n \score_\theta(B_i)
  - \frac{1}{4} \sum_{i = 1}^n W_{n,i}^2 + o_P(1).
\end{equation}

It remains to compute $\E[W_{n,i}^2]$. Using the
expansions~\eqref{eqn:h-fourth}, we have for any $B = \indic{X \in A}$,
%% \begin{align*}
%%   \E\left[
%%     \left(\sqrt{\frac{P_{\theta + h}(B)}{P_\theta(B)}}
%%     - 1\right)^2
%%     \right]
%%   & = (\sqrt{P_{\theta + h}(A)} - \sqrt{P_\theta(A)})^2
%%   + (\sqrt{P_{\theta + h}(A^c)} - \sqrt{P_\theta(A^c)})^2 \\
%%   & = (h \dPtheta(A) / \sqrt{P_\theta(A)}
%%   + O(h^2)) + (h \dPtheta(A^c) / \sqrt{P_\theta(A^c)} + O(h^2))^2 \\
%%   & = h^2 P_\theta(A) \score_\theta(A)^2
%%   + h^2 P_\theta(A^c) \score_\theta(A^c)^2
%%   + O(h^3)
%%   = h^2 \var(\score_\theta(B))
%%   + O(h^3)
%% \end{align*}
%% uniformly in $h$,
%% and
\begin{align*}
  \left(\sqrt{\frac{P_{\theta + h}(B)}{P_\theta(B)}}
  - 1\right)^2
  & = \left(\sqrt{1 + h \score_\theta(B) + O(h^2)} - 1\right)^2
  % = \left(h \score_\theta(B) / 2 + O(h^2)\right)^2
  = h^2 \score_\theta(B)^2 / 4 + O(h^3)
\end{align*}
uniformly in $h$ near 0, as the probabilities $P_\theta(A)$ are bounded from
0 and 1. Thus
\begin{equation*}
  \sum_{i = 1}^n W_{n,i}^2
  = \frac{h_n^2}{n} \sum_{i = 1}^n \score_\theta(B_i)^2
  + O(h_n^2 / \sqrt{n}),
\end{equation*}
giving Lemma~\ref{lemma:lan-bits}.