### 2019-12-23 preliminaries.tex

\section{Consistent Estimation and Off-the-shelf Bounds \label{sec:preliminary}}

We begin our technical treatment by deriving a few
bounds on the efficiency of estimators in
setting~\eqref{item:distributed}. These bounds establish the
following facts:
\begin{enumerate}[1.]
\item A consistent estimator with an asymptotically normal distribution
  always exists in setting~\eqref{item:distributed}, and hence in the
  adaptive settings~\eqref{item:adaptive} and
  (\ref{item:one-step-adaptive}').
\item For the normal distribution, the asymptotic relative
  efficiency~\eqref{eqn:are-def} in the distributed
  setting~\eqref{item:distributed} is at most $3/4$. No estimator can be as
  efficient as the sample mean.
\end{enumerate}

\subsection{Consistent Estimation}
The simplest estimator is simply to invert a quantile. Indeed,
fix $\theta_0 \in \mathbb R$ and define the $i$th message by 
\[
B_i = \mathbf 1_{X_i<\theta_0}, 
\]
where $\mathbf 1_A$ is the indicator of the event $A$. We have
\[
P_n \triangleq\frac{1}{n} \sum_{i=1}^n B_i \overset{a.s.}{\rightarrow} F(\theta_0 - \theta),  
\]
so that 
\begin{equation}
\label{eq:estimator_naive}
{\theta}_n = \theta_0 - F^{-1}\left( P_n \right)
\end{equation}
is a consistent estimator for $\theta$ in the distributed setting of
Figure~\ref{fig:setup}-(iii), where we note that $F$ is invertible over the
support of $f$. As the variance of $P_n$ is
$F(\theta_0-\theta)\left(1-F(\theta_0-\theta)\right)$, a delta method
calculation~\cite[Ch.~23]{VanDerVaart98} implies that ${\theta}_n$ is
asymptotically normal with variance
\begin{equation*}
  \frac{F(\theta_0-\theta)\left(1-F(\theta_0-\theta)\right)}{f^2(\theta_0-\theta)} = \frac{1}{
\eta(\theta_0-\theta)}.
\end{equation*}
In the Gaussian case where the $X_i \simiid \normal(\theta, \sigma^2)$, the
ARE of ${\theta}_n$ is $\eta(\theta_0 - \theta)\sigma^2$.

Assumption~\ref{assump:failure_rate} implies that the optimal asymptotic
variance for an estimator of the form~\eqref{eq:estimator_naive} is $1 /
\eta(0)$, the asymptotic of the sample median. Unfortunately, as $\theta$ is
(by definition) \emph{a priori} unknown and $\eta(x)$ monotonically
decreases in $|x|$, this naive estimator $\theta_n$ may be very inefficient
when $\theta$ is far from the initial guess $\theta_0$. As an example, when
$f$ is a the normal density, the ARE of ${\theta}_n$ is less than $0.15$
when $|\theta_0 - \theta| \ge 2\sigma$, and more broadly, $\ARE(\theta_n)$
asymptotes to $|\theta_0| \exp(-\theta_0^2 / 2) / \sqrt{2\pi}$ as $|\theta_0
- \theta|$ gets large.  Yet that $\theta_0 = \theta$ minimizes this
asymptotic variance, and $\eta$ is continuous, is suggestive: if we can use
a suitably good initial estimate $\theta_n\init$ for $\theta$, it is
possible that a one-step adaptive estimator
(recall~(\ref{item:one-step-adaptive}')) may be asymptotically strong, as we
see in Section~\ref{sec:sequential}.

\subsection{Multiterminal Source Coding}
\label{sec:ceo}

A related problem is the CEO problem, which considers the estimation of a
sequence $\theta_1,\theta_2\ldots$, where a noisy version of each $\theta_i$
is available at $n$ terminals. At each terminal $i$, an encoder observes the
$k$ noisy samples
\[
X_{i,j} = \theta_j + Z_{i,j},\qquad j=1,\ldots,k, \qquad i = 1,\ldots,n,
\]
and transmits $r_i k$ bits to a central estimator~\cite{berger1996ceo}. The
central estimator produces an estimates ${\theta}_1,\ldots,{\theta}_k$ with
the goal of minimizing the quadratic risk:
\[
R_{\CEO} = \frac{1}{k} \sum_{j=1}^k \mathbb E \left[\left(\theta_j - {\theta} \right)^2 \right]. 
\]
Note that any distributed encoding scheme using one-bit per sample can be replicated $k$ times and thus leads to a legitimate encoding and estimation scheme for the CEO problem with $r_1=\ldots=r_n = 1$. It follows that, assuming that $\theta$ is drawn once from the prior $\pi$, our mean estimation problem from one-bit samples under distributed encoding corresponds to the CEO setting with $k=1$ realization of $\theta$ observed under noise at $n$ different locations, and communicated at each location using an encoder sending a single bit. 
%
Consequently, a lower bound on the MSE in estimating $\theta$ in the
distributed encoding setting is given by the minimal MSE in the CEO setting
as $k \to \infty$. Note that the difference between the CEO setting and ours
lays in the privilege of each of the encoders to describe $k$ realizations
of $\theta$ using $k$ bits with MSE averaged over these realizations, rather
than a single realization using a single bit in ours.

When the prior on $\theta$ and the noise corrupting it at each location are
Gaussian, Prabhakaran et al.~\cite{prabhakaran2004rate} characterize the
optimal encoding and its asymptotic risk as $k \to \infty$.  Chen et
al.~\cite{chen2004upper} also provide an expression for the quadratic risk
in the CEO setting under Gaussian priors. Adapting to our setting, this
expression provides the following proposition:
\begin{prop} \label{prop:CEO}
  Assume that $\Theta = \mathbb R$ and $\pi(\theta) =
  \Ncal(0,\sigma_\theta^2)$ where $\sigma_\theta^2 \in \mathbb R$ is
  arbitrary. Then any estimator ${\theta}_n$ of $\theta$ in the distributed
  setting satisfies
  \begin{equation} \label{eq:ceo_bound}
    n \ex{\left( \theta - \theta_n \right)^2} \geq \frac{4}{3} \sigma^2 + O(n^{-1}),
  \end{equation}
  where the expectation is with respect to $\theta$ and $X_1,\ldots,X_n$.
\end{prop}
\noindent
See Appendix~\ref{app:proof:CEO} for a proof.

As we shall see, this bound is loose: the difference between the MSE lower
bound~\eqref{eq:ceo_bound} and the actual MSE in the distributed setting
(case~\eqref{item:distributed}) occurs because in the CEO setting, each
encoder may encode an arbitrary number of $k$ independent realizations of
$\theta$ using $k$ bits; in our situation, $k = 1$. That blocking allows
more efficient encoding and exploiting the high-dimensional geometry of the
product probability space in the CEO problem is perhaps unsurprising, and
our goal in the sequel will be to characterize the performance degradation
one bit encoding engenders.