### 2021-09-08 abstract.tex
fixed all minor comments from R1

% -*- Mode: latex -*- %
\begin{abstract}
  We consider the problem of estimating the mean of a symmetric log-concave
  distribution under the constraint that only a single bit per sample
  from this distribution is available to the estimator. We study the mean
  squared error as a function of the
  sample size (and hence the number of bits).
  We consider three settings: first, a centralized setting, where
  an encoder may release $n$ bits given a sample of size $n$, and for
  which there is no asymptotic penalty for quantization; second, an adaptive
  setting  in which each bit is a function of the current
  observation and previously recorded bits, where we show that
  the optimal relative
  efficiency compared to the sample mean is precisely
  the efficiency of the median; lastly, we show that in
  a distributed setting where each bit is only a function
  of a local sample, no estimator can achieve optimal efficiency
  uniformly over the parameter space.
  We additionally complement our results in the adaptive setting
  by showing that \emph{one} round of adaptivity is sufficient
  to achieve optimal mean-square error.
  %% single sample. We show that the optimal efficiency of the adaptive
  %% setting can be attained by splitting the sample into two parts and
  %% allowing only a single adaptation. Otherwise, we show that no one sample
  %% estimation procedure can attain the optimal efficiency uniformly over all
  %% points in the parameter space.  Our results indicate that estimating the
  %% mean from one-bit measurements is equivalent to estimating the sample
  %% median from these measurements. In the adaptive case, this estimate can be
  %% done with vanishing error for any point in the parameter space. In the
  %% distributed case, this estimate can be done with vanishing error only for
  %% a finite number of possible values for the unknown mean.
\end{abstract}