### 2022-04-20 ak-bio.tex

Alon Kipnis is a Senior Lecturer at the Efi Arazi School of Computer Science at Reichman University. Preivously, he was a postdoctoral research scholar and a lecturer at the Department of Statistics at Stanford University, advised by David Donoho. He completed the Ph.D. in electrical engineering from Stanford University in 2017. Dr. Kipnis is the recipient of numerous fellowships and awards, including the Koret foundation postdoctoral fellowship, Stanford's electrical engineering department fellowship, and a prize from the Wolf Foundation. His research is in the areas of mathematical statistics and information theory.